### 2006-04-05 TTau_Halpha/figures/appedix.tex

%% LyX 1.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{graphicx}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

\usepackage{babel}
\makeatother
\begin{document}
\begin{tabular}{c}
\includegraphics[%
  scale=0.45]{/usr/local/home/rk/paper/TTau_Halpha/figures/discwind_hybrid_mdot7_inc10.eps}\tabularnewline
\includegraphics[%
  scale=0.45]{/usr/local/home/rk/paper/TTau_Halpha/figures/discwind_hybrid_mdot7_inc55.eps}\tabularnewline
\includegraphics[%
  scale=0.45]{/usr/local/home/rk/paper/TTau_Halpha/figures/discwind_hybrid_mdot7_inc80.eps}\tabularnewline
\end{tabular}
\end{document}