### 2006-04-05 TTau_Halpha/revision01/response_to_ref.txt
tims changes
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@695 25464f21-6099-402c-b30a-640f1d40d38b

Dear Referee,

Thank you very much for reviewing our paper, and for doing so in
timely mannar.  We have reviewed our paper by considering your
comments and suggestions very carefully.  
The answers to your questions follows "*" marks before 
the indented texts below.

Kind regards,

Ryuichi Kurosawa, Tim Harries and Neil Symington

===================================================================
General comments

Introduction, 3rd paragraph: The blue-ward asymmetric emission is
generated in magnetospheric accretion models not only by the
occultation of the flow by the stellar photosphere, which occults
mostly high velocity blue-shifted material, but also by the
occultation of part of the receding flow by the inner disk, which
occults mostly low velocity redshifted emission. Both effects (lack of
low velocity red-shifted emission and lack of high velocity
blue-shifted emission) contribute to generate the blue-ward asymmetry.

   * Thank you for point that out. I replaced the sentence 
     from 
	  "... blue-ward asymmetric emission line 
	    profiles as resulting from the partial occultation 
	    of the flow by the stellar photosphere, .... 
     to

	  "... blue-ward asymmetric emission line 
	    profiles as resulting from the partial occultation 
	    of the flow by the inner part of the accretion disc and by
	    the stellar photosphere, .... "

Section 2.1, 1st paragraph: The adopted parameters of a typical
classical T Tauri star are listed in Table 1. I think that a mass
accretion rate of 10^-8 Msun/yr is much more representative of a
typical CTTS than 10^-7 Msun/yr. The values you adopted as typical for
both Macc and Mwind rates are actually quite high compared to what is
usually measured (see for ex. Gullbring et al. 1998). It is true that
some CTTSs present such high Macc and Mwind rate values, but these are
only a few. This is just a comment and does not invalidate your
results since you ended up calculating profiles for Macc = 10^-7,
10^-8 and 10^-9 Msun/yr, which clearly encompasses the whole range of
measured Macc rates in CTTSs.

    * I agree. A typical value of mass-accretion rate is ~3x10^-8
      Msun/yr according to Gullbring et al. (1998) and  Calvet et al. (2004).
      A slightly higher mass-accretion was used so that the line 
      broadening can be seen more clearly, and because 
      10^-7 is a nice round number. I clarify this point in caption of
      Table 1.

Section 2.1, 2nd paragraph: A better reference to the accretion schock
characteristics would be Calvet & Gullbring (1998) ApJ 509, 802 and
Gullbring et al. (2000) ApJ 544, 927, instead of Konigl (1991) and
Hartmann et al. (1994).

     * I replaced the references as you suggested.

In this paragraph you obtain about 8% of the stellar surface covered
by hot rings. This is quite a large value compared with the values
obtained by Gullbring & Calvet in the papers cited above, which range
rather from 0.1% to 1% for several CTTSs. But this could be due to the
fact that you are using a large Macc rate (10^-7 Msun/yr) for a
CTTS. Does this affect your results ? Was the hot ring surface
coverage recalculated consistently when you computed models with 10^-8
Msun/yr and 10^-9 Msun/yr ? In that case, what values were used ?

   * The surface "area" covered by the hot rings depends only on 
     the geometry of the magnetosphere (Rmi and Rmo); however, 
     the total accretion luminosity (L_acc) depends on the
     mass-accretion rate (Macc).
     The surface intensity of the hot rings is computed according to 
     this total luminosity due to accretion and by assuming they
     radiates as a blackbody.

     If a smaller hot ring surface was adopted, the ring temperature
     will increase in order to have a fixed L_acc for a given amount
     of mass-accretion rate.  So, yes, we calculate the amount of
     total "flux" according to a mass-accretion rate, but the surface
     coverage does not change.

     The level populations are relatively insensitive to the actual 
     distribution of the hot spots on the stellar surface and
     hence, the H-alpha line strength is not affected by the change in
     the area coverage of the rings provided that the change is not too
     greate. If the area coverage is much bigger (e.g. 30%), then the
     geometry of the magnetosphere itself must change significantly.  This
     would result in the line strength slightly for a fixed
     inclination angle, but overall shape of the profile would not
     change so much. For example, we used Rmi=2.2R* and Rmo = 3.0R*
     (Small and Wide (SW) model of Muzerolle et al 2001) which results
     in 8% hot ring coverage. If we were to use Rmi=2.2R* and Rmi =
     2.8R* and Rmo=3.0* (Small and narrow (SN) model of Muzerolle et
     al 2001), then the coverage would be 1% of the total surface,
     which is more similar to the fraction estimated in Gullbring et al
     (2000).  If we look at the model comparison of Muzerolle et al
     (2001) (see their Fig. 5), the line strength changes, but the
     line profiles shape do not change much (except for the 10^3 K
     model). We confirmed, by running additional models, the line
     shapes do not change significantly for H-alpha by changing from
     SW to SN model.

     To make it clear about this point, I have inserted the following
     texts near the end of the 2nd paragraph in  Section 2.1:

        "In our model, the fraction of hot-ring coverage area is
        determined only by the geometry of the magnetosphere (Section
        2.2). Gullbring et al (2000) estimated the fractions of the
        hot-ring coverage area for three intermediate CTTS by using
        the observed UV and optical spectra. They found the range of
        fraction to be 0.3--3.0~per~cent.  This is smaller than the
        value adopted here; however, even if we use 1~per~cent
        hot-ring area coverage model (by changing the width of the
        magnetosphere), we find the line shapes of H$\alpha$ do not
        change significantly (see also figure~5 of Muzerolle et al. 2001). "

     Also I added the following sentence, just before the last
     sentence in Section 2.1:

        "The luminosity ratio and the temperature depend on a adopted
         mass-accretion rate."

Sect. 2.2, 3rd paragraph: You say that the results of Martin (1996)
qualitatively agree with those of Hartmann et al (1994). There is a
comment about that in the Muzerolle et al (1998) (ApJ 492, 743) paper
on page 751, where they also use the Hartmann et al (1994) law. They
say that the Martin (1996) temperature distribution is inadequate to
obtain line emission consistent with the observations, even at Halpha,
while their temperature distribution generates profiles that look like
the observed ones. So I think that you should not say that both
temperatures agree, even qualitatively.  It gives the idea that they
generate the same types of profiles, which is not true. The fact that
they do not agree is actually a big problem, since the Hartmann et
al. (1994) temperature structure, which is somewhat arbitrary,
reproduces the observed profiles, while the Martin (1996) temperature
structure, which was calculated consistently, does not. We are missing
some physics in there.

   * The two temperature structures _do_ agree in a qualitative sense,
     in that their behaviour as a function of distance along a
     streamline follows a similar dependence. Nonetheless, you are
     right that Muzerolle could not reconcile the Martin T structure
     with observations. We have made the following change:

        "Muzerolle et al (1998) found that the line profile models computed
         according to the Martin (1996) temperature structure do not
         agree with observations, unlike profiles based on the (less self-consistent)
         Hartmann et al. (1994) temperature distribution.
	 It is clear that The temperature structure of the magnetosphere is still a large
         source of uncertainty in the accretion model; here  we adopted the model of Hartmann et
         al. (1994). "

Sect. 2.3, 4th paragraph: I could not follow you in the end of the
paragraph. You have p=4*alpha*q and q = -1.15 from Whitney et
al. (2003). You say that in order to be consistent with Krasnopolsky
et al. (2003) who used p = -5/2, the value of alpha was set to
0.76. Well, alpha=p/(4*q)=-2.5/(4*(-1.15))=0.54 and not 0.76.

   * You are absolutely correct. This is a typo, 
     it should be p=-7/2 which is the value used in the models.  
     With p=-7/2,  
         alpha=p/(4*q)=-3.5/(4*(-1.15)) = 0.76.

	 ==> Corrected.

Sect. 4.1, 1st paragraph Tmax is not defined. I know it is the maximum
temperature of the accretion flow, but you must define it.

   * I added the following sentences after the first sentence of the
     same paragraph:
       "The temperature structure of Hartmann et al. (1994, ApJ, 426,
	669) is  adopted here, and the parameter Tmax sets
	the maximum temperature of the accretion streams. "

Sect. 4.1, 5th paragraph You say the EWs decrease as the inclination
increases mainly because the fraction of the accretion stream blocked
by the photosphere increases. I agree that the fraction of the
accretion stream blocked by the photosphere increases a lot as the
inclination increases (in Fig. 1, lower left funnel portion or z<0 and
x<0 for an observer looking from the upper right corner), but the
fraction blocked by the inner disk (at z<0 and x>0 in Fig. 1) also
increases a lot and contributes to the decrease in the EWs. The
photosphere mostly blocks blue-shifted material, while the inner disk
mostly blocks redshifted material.

   * I agree. I changed the sentence as you pointed
     out.

     ORIGINAL 
       "The line flux and the EWs decrease mainly because the fraction
        of the accretion stream blocked by the photosphere increases
        as the inclination increases."

     REVISED
       "The line flux and the EWs decrease mainly because (1) the
        fraction of the accretion stream blocked by the photosphere
        increases as the inclination increases (c.f. Fig. 1,
        lower-left part of the accretion stream i.e.  z<0 and x<0),
        and (2) the fraction of the accretion stream blocked by the
        inner part of the accretion disc increases as the inclination
        increases (c.f. Fig. 1, lower-right part of accretion stream
        i.e. z<0 and x>0). While the photosphere mostly blocks the
        blue-shifted part of the stream, the inner disc blocks the
        red-shifted part."

Sect. 4.1, 6th paragraph I don't understand why the highest
red-shifted line-of-sight velocity is visible only at the high
inclination angles. Looking at Figure 1, at low inclinations (even at
i=0) we actually see all the highest red-shifted material of the upper
right funnel portion (z>0, x>0, for an observer looking from the upper
right corner). Actually, looking at the profiles in Fig. 5, I don't
see much of a change in the redshifted wing width (for example at v >
120 km/s) with the increase in inclination. The blue-shifted wing
changes more.

   * Sorry. I meant to say "the blue-shifted" line-of-sight velocity
     instead of the red-shifted ...  As you mentioned the difference
     in the red wing is very small, but the difference in the blue
     wing is small but visible. By reflecting these points, I have
     made the correction as follows:

       ORIGINAL
       "..., the highest red-shifted line-of-sight velocity is visible
        only at the high inclination angles. This explains the wider
        appearance of the profile with i=80 deg. compared to the
        relatively narrow line appearance of the profile with
        i=10 deg."

       REVISED 
       "..., the highest blue-shifted line-of-sight velocity is seen
        only at the high inclination angles. For example, the
        accretion stream close to the photosphere (~1 R*) above the
        photosphere) on the upper-left quadrant in Fig.1 (x<0 and z>0)
        will be more aligned with the line-of-sight of an observer in
        the upper-right quadrant in the same figure (x>0 and z>0) as
        the inclination increases.  This results in a slight increase
        in the blue wing flux for a higher inclination model as seen
        in Fig. 5."

   * In addition to this correction, the last sentence in this
     paragraph (Sect. 4.1, 6th paragraph),

     "A similar EW dependence on
      inclination angle is found for the models with different 
      magnetospheric temperatures and the mass-accretion rates."

      was moved to the end of the previous paragraph (Sect. 4.1, 6th
      paragraph) because it was out of place.

Sect. 4.2, end of 2nd paragraph: You comment that the blue edge of the
absorption component in the slowest wind acceleration model (beta=4.0)
is located close to line centre (~20 km/s). But you don't show
beta=4.0 models. Was that a mistake and you intended to say beta=2.0 ?
I think you should cite the blue edge absorption component velocity of
the fastest model shown in Fig.6, which is beta=2.0. It is easier for
the reader to be able to check in the figure what you are explaining
in the text, and you make your point in the same way.

   * I meant to say beta=2.0 instead of
     beta=4.0.  ==> Corrected.

Sect. 4.2, Table 4: Were does the value of p= -3/2 come from ? In
Section 2.3 you said you would be consistent with the model of
Krasnopolsky (2003) who used p= -5/2.

   * It should be p= -7/2.  
     A similar correction was made in Section 2.3 as you also pointed out.
     ==> Corrected.

Sect. 4.2, 4th paragraph: The opening angle is measured from the
symmetry axis (z), then it should be given by theta_open=arctan(Rwi/d)
instead of theta_open=arctan(d/Rwi). High values of d correspond to
more collimated winds and therefore to smaller opening angles.

   * This was a typo. It should have written as
     theta_open=arctan(Rwi/d). 
     ==> Corrected.

Sect. 4.2, end of 4th paragraph: You say that you used d=44 Rstar for
the disc wind models presented in the paper but Table 4 has d= 22
Rstar listed. Which one is correct ? If Table 4 is a summary of the
reference disc wind models used in Section 4.2, maybe you should not
list d=22 Rstar, since you actually varied d in that section. Either
you put the range of values used (from 2 Rstar to 88 Rstar, as in
Fig. 7) or you don't put any value for d in Table 4.

   * This is my mistake again. I got confused between R* and Rsun.
     The largest displacement (d) model is "d=22R*." 
     d=44R* and d=88R*
     should be replaced with d = 22R*. "d=22R*" is the value used in 
     the models shown in Figure 6.
     ==> The values are corrected in the text and in the caption of
     Figure 7.

   * Deleted the entry for d from Table 4 as suggested.

Sect. 4.2: A general comment on the wind models presented in this
section. Blandford & Payne (1982) showed that in order to launch a
wind magneto centrifugally, the poloidal component of the magnetic
field must make an angle of less than 60 degrees with the disk
surface. In order for that to be achieved in your model, you must have
(supposing a flat disc, which should be reasonable for the inner disk
regions):

d < Rw*tg(theta), where theta = 90 - theta_open = 60 degrees is the
angle between the field line and the disk surface (x axis).

This means that for Rw = Rwi = 3 Rstar, you must have d < 5.2 Rstar
and that for Rw = 1 au ~ 100 Rstar, you must have d < 173 Rstar

The last condition is clearly satisfied in all your models, since d is
always less than 173 Rstar. But the inner field lines, passing close
to Rwi, do not satisfy the Blandford & Payne (1982) condition, since
for that you would need to have d < 5.2 Rstar in all cases.

In other words, if you fixed d = 44 Rstar in the disc wind models
presented in the paper, as you say you did in the 4th paragraph of
Sect. 4.2, for the wind to be launched centrifugally according to
Blandford & Payne, the field lines must present Rw > d/tan(60), which
is Rw > 25.4 Rstar. This means that field lines from Rwi = 3 Rstar to
Rw = 25 Rstar will not be able to launch a wind magneto-centrifugally
with d=44 Rstar.

How do you launch the wind in the inner field line region ? If you
cannot launch a wind in that region, how does it affect your results ?

   * I noticed the launching angle (measured from disc plane)
     condition of Blandford & Payne (1982) is not satisfied in the
     inner region of our simple disc wind model.  This is was done so
     that wind has a moderate amount of collimation. Since the stream
     line is straight, the wind will never bend toward the symmetry
     axis (z-axis) as the distance from the disc increases unlike the
     model of Blandford & Payne (1982). We therefore used a larger
     launching angle to simulate the stream lines in the far field (at
     a larger distance from the disc plane) in their model.  This is
     an intrinsic problem with approximating the stream lines with
     straight lines.  Also, in our mode, if
     the launching angle is too small (<60 deg.), the wind absorption
     feature would not been seen for profiles computed with i<60
     deg. since the wind stream line is straight and little material
     is present in the line-of-sight to an observer in this case.
     Since more than 50% of observed H-alpha profiles (e.g. Reipurth
     1996, A&AS, 120, 229 and Alencar & Basri 2000, ApJ 119, 1881) are
     affected by wind absorption, it was necessary for our model to
     used the launching angle (at the inner most part) greater than 60
     degrees to be more consistent with observations.

     If the launching angle (measured from the disc plane) is 60 deg. at
     the inner most part of the  wind, and the inclination angle
     (measured from the symmetry  axis, z-axis) is less than 60 deg., little
     absorption is expected, as demonstrated in Fig. 7.
     The situation will however change if the
     field/stream line is not straight and bending towards the
     symmetry axis as the wind travel higher above the disc.

     In summary, this was necessary compromise in order (1) to be more
     consistent with observation, and (2) to simulate more
     "colliminated" wind (toward the symmetry axis).

     To answer your question how it would effect our results, I ran
     additional models with d=5Rstar which satisfies the wind
     launching condition of Blandford & Payne (1982) in all disc wind
     zone in our model.  These models showed that the wind absorption
     and emission features weakens,  but overall morphology dependence
     of the profiles on the wind temperature and the wind acceleration rate
     is very similar to what we found in the models shown in Figs. 6
     and 9.  Hence, the major conclusions obtained by using the wind
     models with a larger launching angle (at the very inner part of
     the disc) presented in this paper would not differ from the
     conclusions obtained from the disc wind models with d=5Rstar.

     These models (for  Mdot_acc=1x10^-8 Msun/yr & Mdot_wind=1x10^-9
     Msun/yr cases) with d=5Rstar are posted on the following website.

     http://www.astro.ex.ac.uk/people/rk/profiles/extra2

     To clarify the points you raised above, I have inserted the
     following in text (mostly the same as mentioned above).

     After the 4th paragraph in Section 4.3, the following new
     paragraphs are inserted:

       "Blandford & Payne (1982) showed that in order to launch a wind
        from a disc plane magneto-centrifugally, the launching angle
        (theta_L, measured from the equatorial disc plane) must be
        less than 60 degrees. In our models, the wind launching zone
        is restricted in between Rwi < wi < Rwo (see Fig. 1 and Table
        1) with Rwi = 3 Rstar and Rwo=1 au.  This sets the upper limit
        of the source displacement distance (d) to be around 5.2 Rstar
        if we want to satisfy the launching angle condition of
        Blandford & Payne (1982) for all the stream lines in our disc
        wind model.  In the models presented in Fig. 6 (and later in
        Fig. 9 in next section), we have adopted d =22 Rstar
        instead. We have used this larger source
	displacement distance for the following two reasons: (1) to
	simulate a more 'colliminated' wind (toward the symmetry axis,
	and (2) for greater consistency with the fraction of H-alpha
	profiles affected by wind absorption in observations (ref xxxxx).

      	Since the stream lines in our simple disc wind are represented
	by straight lines, the wind never bends toward the symmetry
	axis (z-axis) as the distance from the disc increases unlike
	the model of Blandford & Payne (1982). We therefore used a
	larger launching angle to simulate the stream lines in the far
	field (at a larger distance from the disc plane) in their
	model.  This is an intrinsic problem with approximating the
	stream lines with straight lines.  Also, in
	our mode, if the launching angle is too small (<60 deg.), the
	wind absorption feature would be not been seen or would be
	very small for profiles computed with i<60 deg.  If the
	launching angle is 60 deg. at the inner most part of the wind,
	and the inclination angle (measured from the symmetry axis,
	z-axis) is less than 60 deg., little (wind) absorption is
	expected, as demonstrated in Figure 7.  The situation will
	however changes if the field/stream line is not straight and
	bending towards the symmetry axis as the wind travel higher
	above the disc.  Since more than 50 per cent of observed
	H-alpha profiles (e.g. Reipurth 1996, A&AS, 120, 229 and
	Alencar & Basri 2000, ApJ 119, 1881) are affected by wind
	absorption, it was necessary for our model to used the
	launching angle (at the inner most part) greater than 60
	degrees to be more consistent with observations; although this
	would mean that the launching angle condition of Blandford and
	Payne (1982) is violated for wind from a very inner part of
	the disc.

	 In addition to the models given in Figs. 6 and 9, we have
	 computed the models with d=5Rstar (not shown here) which
	 satisfies the wind launching condition of Blandford & Payne
	 (1982) in all parts of the disc wind zone in our model.
	 These models showed that the wind absorption and emission
	 features weaken, but overall morphology dependence of the
	 profiles on the wind temperature rate and the wind
	 acceleration rate are similar to the models
	 shown in Figs. 6 and 9.  The minimum inclination angle at
	 which the wind absorption becomes important is larger than
	 that of the models with a larger source displacement (d = 22
	 Rstar) as expected."

Sect. 4.3, 1st paragraph: You say the parameters used for the
magnetosphere are the same as in Fig.5, but in that figure you have
models with 3 different inclinations. Did you calculate models for the
three inclinations or just for i=55 degrees ? Please specify this in
the text.

   * All the models in Section 4 are computed at i=55 deg. The
     profiles computed at 10 and 80 deg. are given in Appendix A. 
     I see that is confusing. 
     I corrected the sentence as follows:

     ORIGINAL
        "The parameters used for the magnetosphere are  same as in 
         Fig. 5, ...."

     REVISED
        "The parameters used for the magnetosphere are same as in 
         Fig. 5 (with i=55 deg.), ...."

   * Also the last sentence is the same paragraph changed to reflect
     your comment:

     ORIGINAL
       "Fig. 9 shows the models profiles computed 
        using the same ranges of the disc-wind temperature and wind
        acceleration parameter as in the previous cases."

     REVISED
       "Fig. 9 shows the models profiles computed (at $i=55^\circ$)
        using the same ranges of the disc-wind temperature and wind
        acceleration parameter as in the previous cases."

Sect. 4.3, 5th paragraph: It is hard to see in Fig. 10 that the
position of minimum flux in the absorption moves blue-ward. I actually
don not see any difference between the models with mu=0.1 and 0.2.

   * You are correct.
     I rephrase the sentence as follows

     ORIGINAL 
     "The figure shows that as $\mu$ increases the
     P-Cygni absorption deepens only slightly, and the position of minimum
     flux in the absorption trough appears to move blue-ward."

     REVISED
     "The figure shows that as $\mu$ increases the
     P-Cygni absorption deepens only slightly  and the position of minimum
     flux in the absorption trough appears to remain at same location."

Sect. 5.1, Type III-B: This is the most common profile type in the
Reipurth et al. (1996) sample but you only find it in you Macc = 10^-7
Msun/yr models, which is not a common or average value for Macc rate
in CTTSs. As said in some previous comment, CTTSs tend to have Macc =
10^-8 Msun/yr and only a few present Macc = 10^-7 Msun/yr. To what do
you attribute that discrepancy between the data and the models ?

   * I have run additional models with a typical mass accretion rate
     (3x10^-8 Msun/yr, Gullbring et al. 1998; Calvet et al. 2004) for
     CTTSs.  Please see
     http://www.astro.ex.ac.uk/people/rk/profiles/extra

     for the profile computed at i=10,55,80 and for beta=0.2, 0.5,
     1.0, 2.0 Twind=7000, 8000, 9000, 10000K

      According to these models which use a more typical
      mass-accretion rate, Type III-B can be seen for at least in
      three models with fast accelerating and high wind temperature
      i.e. (beta, Twind) = (0.2, 10000K), (0.5, 9000K) and (0.5,
      10000K).

      I added a following comments based on these additional runs.

      "Although Type III-B profiles do not occur at macc=10^-8, we
      have examined the intermediate case of macc=3x10-8, which may
      represent a typical T Tauri mass accretion rate  (Gullbring et
      al. 1998; Calvet et al. 2004). We find that the models with fast accelerating,
      high temperature winds do display  Type III-B profiles. It is
      clear that our model, as it stands, requires above-average (but
      perhaps not atypical) mass-accretion rates in order to produce
      Type III-B profiles. It is less clear whether the 14 Type-IIIB
      objects found by Reipurth et al. consistute the highest
      mass-accretors in their sample: the aveage EW for T~Tauri stars
      with Type III-B profiles is 69\AA, which is in the top quartile
      of their complete T~Tauri star sample, perhaps indicating that
      Type-IIIB profiles are associated with high EWs, and therefore
      strong mass-accretion."

Sect. 5.1, Type IV-R: In axisymmetric models (rotation and magnetic
axis aligned) the inverse P Cygni profile really restricts the
inclination angle to a small range. But there are observational
evidences (SU Aur: Johns & Basri 1995, AA Tau: Bouvier et al. 2003,
for example) for inclined magnetospheres and in that case you could
see the accretion shock through the fast moving accretion flow at
basically any inclination. It would only depend on a combination of
the rotation axis and magnetic axis inclinations. In Halpha, in
particular, it might still be difficult to see inverse P Cygni
profiles because of the wide emission wings due to broadening effects
(such as the Stark effect).

   * Yes, you are quite right. I am aware of the effect of the
     inclined magneto-axis CTTS models. In fact, I wrote a paper on SU
     Aur Pa-beta line variability. H-alpha variability of of SU Aur
     indeed shows a very complex line profile morphology dramatically
     changing with its rotational phase (e.g.  Johns & Basri 1995, see
     also their movie
     http://sprg.ssl.berkeley.edu/~cmj/html/suaur.mpg).  Although it
     does not look it has a clear IV-R type profile at any given
     rotational phase for H-alpha, I am aware that Pa-beta line does
     show IV-R type profile, and the strength of inverse P-Cyg or the
     absorption features in the redwing changes as the rotational
     phase changes (e.g. Kuorsawa et al, 2005, MNRAS, 358, 671).

     I wanted to restrict the cases for purely axi-symmetric
     geometries in this paper for clarity of discussion.  In reality,
     a situation may not be as simple as this classification, as we
     know the lines are sometimes highly variable. This problem will
     be tackled in near future. For a starter, I am currently
     computing line profiles based on the MHD model of Romanova et al
     (2003, ApJ, 595, 1009) who considered the disc accretion problem
     with an inclined dipole magnetic field, and investigate if the
     overall line variability features seen in the observation can be
     explained by this model. I am working on the case with and
     without the additional outflow (discwind or/and stellar wind).

     I will clarify this point by adding the following near the end of
     the first paragraph in Section 5.1 :

       "Although in general H-alpha in the spectra of CTTS shows a
        significant amount of variablity, which may in some cases
        induced by the rotational motion of an inclined magnetosphere
        (e.g. Johns &  Basri 1995; Bouvier et al. 2003), we restrict
        our discussion in this section to the axisymmetric case only."

    Also a short comment is added at the end of Type IV-R section,

       "The restriction in the inclination angle mentioned above would
        be relaxed if the axis of the magnetosphere was tilted with
        respect to the rotational axis, since the line-of-sight from an
        observer to the hot rings will pass through 
        different parts of the magnetosphere as the rotational phase
        changes."

Sect. 5.2, 3rd paragraph: What is the mass loss rate used in the
stellar wind models ? This value should be added to the text. Matt &
Pudritz (2005) suggest 10^-9 Msun/yr to solve the angular momentum
problem.

   * Thank you for pointing this out.  I should have quoted the
     mass-loss rate used in the models.    
     In both models in Fig.13, Mdot_wind = 10^-8 Msun/yr is used so 
     that mu=Mdot_wind/Mdot_acc = 0.1, and to be consistent with 
     the mass-accretion rate (Mdot_acc=10^-7 Msun/yr) used for the
     models in earlier in the paper. It seems that the ratio mu is an
     important factor in Matt & Pudritz (2005) in balancing the
     torques on a central star (their equation 3).
     I tried the models with Mdot_wind = 10^-9 Msun/yr  (keeping
     mu=0.1), and found the same inclination dependency for these
     models, so the main conclusion of this section 
     does not change when Mdot_wind is changed from 10^-8 Msun/yr to
     10^-9 Msun/yr.

     To reflect your points, I added the following in the text 
     (Sect. 5.2, 4th paragraph).

     After the first sentence of the paragraph, I inserted:

        "The mass-loss rate of the stellar wind in this model is 10^-8
	Msun/yr, which is same as the mass-loss rate used in the
	disc-wind magnetosphere hybrid model in the same figure.  In
	both models, the mass-accretion rate of 10^-7 Msun/yr is
	used."

     At the end of the paragraph, I inserted: 
        "A similar inclination dependency of EW is found for lower
        mass-loss rate models (e.g. Mdot_wind = 10^-9 Msun/yr)
        while keeping mu=0.1."

     Also in the figure caption of Figure 13  just before the last
	sentence,  I added:
	"The wind mass-loss rates used in both models is Mdot_wind =
	10^-8 Msun/yr."

***** Spells and other minor comments

Sect. 2.1, 2nd paragraph: ... and re-emitted optical and UV radiation.
Maybe rephrase as: ... and re-emitted as optical and UV radiation.

   * Corrected.

Sect. 2.2, 1st paragraph: Rmi and Rmo --> 2.2 Rsun and 3.0 Rsun should
be 2.2 Rstar and 3.0 Rstar

   * Corrected.

Sect. 2.3, 5th paragraph: The Rs value is set to 10 Rmi. Shouldn't it
be 10 Rwi, as in Table 4 ?

   * Yes. Thank you for noticing. Changed it to Rwi,

Sect. 2.3, caption of Figure 2: ...the density is approximately
proportional to ~ l^2 It looks like ~ l^-2

   * Yes, you are correct. My mistake again. It should be l^-2.
     ==> corrected.

Sect. 3, last paragraph: ...the dust property described in Section 10. 
¨Section 10¨ should be ¨Section 2.4.2¨

   * Corrected.

Sect. 4.2, caption of Figure 7: The 88 Rstar model is represented by a
solid line and not by a dash-dotted line as written.

   * Corrected.

In the same caption, "opening angel" should be "opening angle"

   * Corrected.

Sect. 4.3, 4th paragraph: "Figures 6" should be "Fig. 6"

   * Corrected.

Sect. 4.3, 5th paragraph: ...the flux peak increase and ...  ¨increase¨ 
should be ¨increases¨

   * Corrected.

Sect. 4.3, caption of Figure 11: The parameter used for the
magnetosphere are....  ¨parameter¨ should be ¨parameters¨

   * Corrected.

Sect. 5.2, 4th paragraph: The sentece below must be rewritten: ...This
tendency (...) holds for the models with this wind models for wide
ranges of beta...

   * Changed the sentence 
     From 
     "This tendency of the inclination dependency holds for the models
      with this wind models for wide ranges of beta and
      T_wind providing that the emission from the wind is
      greater than that of the magnetosphere."

     To
     "This inclination dependency of EW is seen in the models with
      wide ranges of beta and T_wind, provided that the wind emission
      of the model is dominating the emission from magnetosphere."

In addition to the referee's suggestions, we noticed the following 
minor mistakes in the original manuscript, and corrected.

1. The mass-loss rate in figure a1 and figure a2 were miss-labeled in
   the captions.  ==>  corrected.

2. In Figure 13, the velocity ranges out the two plots should be
   same for a better comparison. We changed the range of the 
   right plot to -800km/s <V < 800 km/s.