### 2006-06-12 BD_radvel/tmp.tex

%% LyX 1.4.1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[onecolumn,english,oneside,useAMS, usenatbib]{mn2e}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\setcounter{tocdepth}{3}
\usepackage{graphicx}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Bold symbol macro for standard LaTeX users
\providecommand{\boldsymbol}[1]{\mbox{\boldmath $#1$}}

%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
%% LyX 1.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Bold symbol macro for standard LaTeX users

%% Because html converters don't know tabularnewline

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{times}

\newcommand{\aap}{A\&A}
\newcommand{\aaps}{A\&AS}
\newcommand{\apj}{ApJ}
\newcommand{\apjl}{\apj}
\newcommand{\pasp}{PASP}
\newcommand{\aj}{AJ}
\newcommand{\mnras}{MNRAS}
\newcommand{\apjs}{ApJS}
\newcommand{\aapr}{A\&AR}
\newcommand{\pasj}{PASJ}
\newcommand{\nat}{Nat}

\newcommand{\kmps}{\mathrm{km~s^{-1}}}
\newcommand{\ion}[2]{#1$\,${\sc {#2}}}   % ion, i.e., CII = \ion{C}{ii}
\newcommand{\Kelvin}{\mathrm{K}}
\newcommand{\Msun}{\mathrm{M_{\odot}}}

\makeatother

\usepackage{babel}
\makeatother
\begin{document}

\title{{[}}

\maketitle
Radial and rotational velocities of young BDs and VLM stars]{Radial
and rotational velocities of young brown dwarfs and very low-mass
stars in the Upper Scorpius OB association and the $\rho$~Ophiuchi
cloud core}

\author{{[}}

R. Kurosawa et\,al.]{Ryuichi Kurosawa$^{1}$\thanks{E-mail:rk@astro.ex.ac.uk},
Tim J. Harries$^{1}$ and S.~P.~Littlefair$^{2}$\\
 $^{1}$School of Physics, University of Exeter, Stocker Road, Exeter
EX4~4QL. \\
 $^{2}$Department of Physics and Astronomy, University of Sheffield,
Sheffield S3 7RH}

\date{Dates to be inserted}

\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2006}

\label{firstpage}

\begin{abstract}

We present the results of a radial velocity (RV) survey of 14 brown
dwarfs (BDs) and very low-mass (VLM) stars in the Upper Scorpius OB
association (UScoOB) and 3 BD candidates in the $\rho$~Ophiuchi
dark cloud core. We obtained high-resolution echelle spectra at the
VLT using Ultraviolet and Visual Echelle Spectrograph (UVES) at two
different epochs for each object, and measured the shifts in their
RVs to identify candidates for binary/multiple systems in the sample.
The average time separation of the RV measurements is 21.6~d, and
our survey is sensitive to the binaries with separation $<0.1$~au.
We found 4 out of 17 objects (or $24_{-13}^{+16}$~per~cent by fraction)
show a significant RV change in 4--33~d time scale, and are considered
as binary/multiple `candidates.' We found no double-lined spectroscopic
binaries in our sample, based on the shape of cross-correlation curves.
The RV dispersion of the objects in UScoOB is found to be very similar
to that of the BD and VLM stars in Chamaeleon~I (Cha~). We also
found the distribution of the mean rotational velocities ($v\,\sin i$)
of the UScoOB objects is similar to that of the Cha~I, but the dispersion
of $v\,\sin i$ is much larger than that of the Cha~I objects.

\end{abstract}

\begin{keywords} stars: binaries:spectroscopic -- stars: low-mass,
brown dwarfs -- stars: formation -- stars:planetary system: formation
\end{keywords}

\section{Introduction }

\label{sec:Introduction}

Most stars are member of binary systems and it is therefore important
that a complete star formation theory be able to predict the binary
fraction, period distribution, and mass-ratio distribution of stellar
objects across a wide range of masses. Furthermore, the study of individual
binary systems is the only direct means to determine fundamental stellar
properties such as stellar masses and radii.

Recent high-resolution imaging studies of brown dwarfs (BDs) and very
low-mass (VLM) stars have placed strong constraints on binaries with
separations of $\sim1-100\,\mathrm{au}$. For example, Hubble Space
Telescope (\emph{HST}) observations of $\alpha$~Per and the Pleiades
indicates a binary fraction ($f$) of $>10$~per~cent with a bias
towards separations ($a$) of less than $15$~au, and a mass-ratio
($q$) of $>0.7$ \citep{martin:2003} for objects around and below
the hydrogen burning limit (see also \citealt{bouy:2006}). A similar
lack of wide binaries was found in the field T-dwarf study \citep{burgasser:2003},
while $f\approx15$~per~cent (for the field objects with 0.03--0.1~$\mathrm{M_{\sun}}$
and M8.0--L0.5) was determined by \citet{close:2003} using the adaptive
optics at Gemini North. They also found the vast majority of systems
have a semimajor axis $<20\,\mathrm{au}$ (see also \citealt{siegler:2005}
but \citealt{luhman:2004}). An \emph{HST} study of more than 80 field
late M and L dwarfs \citep{gizis:2003} indicated $f\approx15$~per~cent
with separations in the range of 1.6--16~au. For a small (12) sample
of BDs and VLM stars (0.04--0.1~$\mathrm{M_{\odot}}$) in Upper Scorpius
OB association (UScoOB), \citet{kraus:2005} found $f=25_{-8}^{+16}$~per~cent
for $5\,\mathrm{au}<a<18\,\mathrm{au}$ by using a similar imaging
technique. More recently, \citet{basri:2006}, combined with the results
of earlier works, found the upper limit of the overall binary fraction
for VLM stars of $26\pm10$~per~cent.

Using a Monte Carlo simulation, the data from radial velocity surveys
available in the literature, and by carefully considering the sensitivity
and sampling biases, \citet{maxted:2005} found an overall BD/VLM
binary frequency of 32--45 per~cent assuming $f=15$~per~cent for
$a>2.6$~au. A recent photometric study \citep{pinfield:2003} of
low-mass objects in Pleiades and Praesepe suggested, albeit indirectly,
$f$ as large as 50~per~cent. which would only be compatible with
direct imaging studies if 70--80~per~cent of those binaries have
$a<1\,\mathrm{au}$. For a more comprehensive review of the current
status of BD/VLM binary fraction and the separation distribution,
readers are refer to a recent review of multiplicity studies by \cite{burgasser:2006}.

The extensive imaging surveys provide excellent observational constraints
on wider BD+BD binaries, but it is now necessary to search for shorter
period BD+BD binaries systematically. Binaries with the separation
of less than 1~au are not resolved by current imaging techniques,
but will be detectable as spectroscopic binaries, providing the mass-ratio
is not too extreme, and velocity separation is large enough. The first
BD+BD spectroscopic binary, PPl~15 \citep{basri:1999}, showed a
double-peaked cross-correlation function with a maximum velocity separation
of $>70\,\kmps$. The binary was found to have an eccentric orbit
($e=0.4$) with a period of $\sim5.8\,\mathrm{d}$. \citet{basri:1999}
suggested that the formation process of substellar objects is biased
towards smaller separation binaries based on the short period of PPl~15
and the lack of Pleiades BD binaries with separations $>40\,\mathrm{au}$.
Note that the median separation of binaries with solar-type primaries
is 30~au \citep{duquennoy:1991}. Pioneering works on the RVs of
BDs and VLMs are presented by \citet{guenther:2003}, \cite{kenyon:2005}
and \citet{joergens:2006} who found a several binary candidates;
however, the orbital parameters and masses of binaries remains unknown
because the follow-up spectroscopic monitoring is lacking or still
being undertaken. In addition to the follow-up observations, the number
of BDs and VLM star binary candidates needs to be increased in order
to have better statistics on short-period binary parameters.

The first BD+BD eclipsing binary (2MASS J0532184--J0546085) was discovered
by \citet*{stassun:2006} from the $I$-band photometric monitoring
of the system. Combining their light curves and the results from the
follow-up radial velocity measurements, they were able to determined
the precise oribital and physical parameters of the system. The projected
semimajor axis and the period of the binary are found as $0.0398\pm0.0010$~au
and $9.779621\pm0.000042$~d respectively.

The separation distribution of BD/VLM binaries is critical to understanding
their origin. There are two main models for the formation of BDs and
VLM stars: first, they have low masses because they form in low-mass,
dense molecular cloud cores (e.g.~\citealt{padoan:2002}); second,
BD/VLM objects have low masses because they are ejected from the dense
core in which they form via dynamical interactions in multiple system,
cutting off their accretion before they have reached stellar masses
(\citealt{reipurth:2001}; \citealt*{bate:2002}). Alternatively,
there is a third model in which a free-floating BD or planetary-mass
object can be formed in the process of the photo-evaporation (e.g.~\citealt{mccaughrean:2002};~\citealt{whitworth:2004})
with the outer layers of a pre-stellar core ($\sim0.2\,\mathrm{M_{\odot}}$)
removed by the strong radiation pressure from the nearby massive OB
stars before the accretion onto the protostar at core centre occurs.

Due to the dynamical interaction involved in the second model, BD/VLM
binaries that survive are generally expected to have small separations.
In the first model, wider binaries may be expected to be more common.
\citet*{bate:2002b} suggested that close binaries ($a<10\,\mathrm{au}$)
do not form directly, but result from hardening of wider systems though
a combination of dynamical interactions, accretion and interactions
with circum-binary discs. If BD/VLM binaries have formed through such
mechanisms, one would not expect to find binaries with 1--10~au separations
without also finding many with separation $<1$~au. If an absence/rarity
of binaries with 1~au were found, it may support the idea that they
are ejected quickly from multiple systems before they have undergone
the interactions that shorten their periods.

Our immediate aim is to identify spectroscopic and close BD/VLM binaries
using the high-resolution echelle spectroscopy at two epochs. This
experiment is sensitive to VLM binaries with separations of $<0.1$~au
which corresponds to a period of $\sim10\,\mathrm{d}$. A larger sample
of candidates will enable us to measure the binary fraction of these
short-period/close binaries (once confirmed), and address whether
there is a significant population of 'hidden' VLM companions. The
long term goal of this project to follow up the binary candidates
found in this paper by spectroscopically monitoring them over different
time scales, enabling us to obtain the radial velocity curves and
their minimum masses.

In Section~\ref{sec:observation-reduction}. we describe the observations
and the data reduction. The results of radial velocity and rotational
velocity ($v\,\sin i$) measurements are presented in Section~\ref{sec:Results}.
We discuss the binary/multiplicity fraction indicated by our RV survey,
and compare our results with earlier works in Section~\ref{sec:Binary-fraction},
and give our conclusions in Section~\ref{sec:Conclusions}.

\section{Observations}

\label{sec:observation-reduction}

Our sample consists of 18 young, very low-mass objects: 15 objects
in the Upper Scorpius OB association ($\mathrm{d\approx145\, pc}$,
\citealt{dezeeuw:1999}) from the list of \citet{ardila:2000} and
3 objects in the $\rho$~Ophiuchi cloud core ($\mathrm{d\approx150\, pc}$,
\citealt{dezeeuw:1997}) from \citet{luhman:1999}. The spectral type
of the objects range between M5 and M8.5, and the age $<\sim10\,\mathrm{Myr}$
(\citealt{luhman:1999}; \citealt{ardila:2000}; \citealt{muzerolle:2003};
\citealt{kraus:2005}). The sample is not complete, and the selection
was solely based on brightness and the observability. The basic properties
of the targets based on the literature is summarised in Table~\ref{tab:literatures}.

We obtained high-resolution spectra with the Kueyen telescope of VLT
(Cerro Parnal, Chile) using the UVES echelle spectrograph. The observations
were carried out between 2004 April 5 and 2004 May 17 in the service
mode. For each object, spectra were obtained at two different epochs
separated by 4--33~d. For each object at a given night, two separate
spectra are obtained consecutively. This allows us to derive more
reliable uncertainty estimates in the RV values of our targets (c.f.~\citealt{joergens:2006}).
The data were obtained using the red arm of UVES spectrograph with
two mosaic CCDs (EEV + MIT/LL with 2k$\times$4k pixels). The wavelength
coverage of 6708 -- 10,250~\AA\, and the spectral resolution $R\approx40,000$
were used. The slit width and length of 1'' and 12'' were used respectively
with a typical seeing of 0.8''.

The data were reduced via the standard ESO pipeline procedures for
UVES echelle spectra. In summary, the data were corrected for bias,
interorder background, sky background, sky emission lines and cosmic
ray hits. They were then flattened, optimally extracted, and finally
the different orders were merged. No binning was performed to achieve
high resolution required for the RV measurements. The wavelength was
calibrated using the Thorium-Argon arc spectra with a typical value
of the standard deviation of the dispersion solution of 5~m\AA\,
which corresponds to 0.2~$\kmps$ at the central wavelength 8600~\AA.
However, the autoguiding of the telescope keeps the star at the centre
of the slit with about a tenth of the FWHM ($1\,\kmps$) which sets
the upper limit for the systematic error in the RV measurements \citep{bailer-jones:2004}.
The accuracy of the wavelength calibration will be demonstrated in
the RV measurements of a RV standard in the following section. A typical
signal-to-noise ratio (S/N) per wavelength bin of the spectra is about
15, and the heliocentric velocity correction was applied to the final
spectra.

%==========================================

\begin{table*}

\caption{Summary of known properties of the targets from literatures: a.~\citet{luhman:1999}
(original list for $\rho$~Oph), b.~\citet{ardila:2000} (original
list for UScoOB), c.~\citet{wilking:1999}, d.~\citet{muzerolle:2003},
e.~\citet{kraus:2005}, and f.~\citet{mohanty:2005}. }

\label{tab:literatures}

\begin{center}\begin{tabular}{llcccc}
\hline 
Object&
Sp.&
mass&
RV&
$v\sin i$&
Known multiple?\tabularnewline
&
&
$\left[\mathrm{M_{\odot}}\right]$&
$\left[\mathrm{km\, s^{-1}}\right]$&
$\left[\mathrm{km\, s^{-1}}\right]$&
\tabularnewline
\hline 
GY 5&
$\mathrm{M7^{c}}$&
$0.07^{\mathrm{d}}$&
$-6.3\pm1.9^{\mathrm{d}}$&
$16.8\pm2.7^{\mathrm{d}}$&
no\tabularnewline
GY 141&
$\mathrm{M8.5^{a}}$&
$0.02^{\mathrm{d}}$&
\ldots{}&
$6.0^{f}$&
no\tabularnewline
GY 310&
$\mathrm{M8.5^{c}}$&
$0.08^{\mathrm{a,d}}$&
\ldots{}&
$10.0^{f}$&
no\tabularnewline
USco 40&
$\mathrm{M5^{b}}$&
$0.1^{\mathrm{b}}$&
\ldots{}&
$37.5^{f}$&
no\tabularnewline
USco 53&
$\mathrm{M5^{b}}$&
$0.1^{\mathrm{b}}$&
\ldots{}&
$45.0^{f}$&
no\tabularnewline
USco 55&
$\mathrm{M5.5^{b}}$&
$0.10+0.07^{\mathrm{e}}$&
\ldots{}&
$12.0^{f}$&
$\mathrm{visual^{e}}$\tabularnewline
USco 66&
$\mathrm{M6^{b}}$&
$0.07+0.07^{\mathrm{e}}$&
$-4.4\pm0.6^{\mathrm{d}}$&
$27.5^{f}$&
$\mathrm{visual^{e}}$\tabularnewline
USco 67&
$\mathrm{M5.5^{b}}$&
$0.10^{\mathrm{e}}$&
\ldots{}&
$18.0^{f}$&
no\tabularnewline
USco 75&
$\mathrm{M6^{b}}$&
$0.07^{\mathrm{e}}$&
$-5.6\pm1.1^{\mathrm{d}}$&
$63.0^{f}$&
no\tabularnewline
USco 100&
$\mathrm{M7^{b}}$&
$0.05^{\mathrm{e}}$&
$-8.9\pm0.6^{\mathrm{d}}$&
$50.0^{f}$&
no\tabularnewline
USco 101&
$\mathrm{M5^{b}}$&
$0.05^{\mathrm{b}}$&
\ldots{}&
\ldots{}&
no\tabularnewline
USco 104&
$\mathrm{M5^{b}}$&
$0.05^{\mathrm{b}}$&
\ldots{}&
$16.0^{f}$&
no\tabularnewline
USco 109&
$\mathrm{M6^{b}}$&
$0.07+0.04^{\mathrm{e}}$&
$-3.8\pm0.7^{\mathrm{d}}$&
$6.0^{f}$&
$\mathrm{visual^{e}}$\tabularnewline
USco 112&
$\mathrm{M5.5^{b}}$&
$0.1^{\mathrm{e}}$&
\ldots{}&
$8.0^{f}$&
no\tabularnewline
USco 121&
$\mathrm{M6^{b}}$&
$0.02^{\mathrm{b}}$&
$-38.9\pm1.0^{\mathrm{d}}$&
\ldots{}&
no\tabularnewline
USco 128&
$\mathrm{M7^{b}}$&
$0.05^{\mathrm{e}}$&
$-3.0\pm1.6^{\mathrm{d}}$&
$0.0^{f}$&
no\tabularnewline
USco 130&
$\mathrm{M7.5^{e}}$&
$0.04^{\mathrm{e}}$&
\ldots{}&
$14.0^{f}$&
no\tabularnewline
USco 132&
$\mathrm{M7^{b}}$&
$0.05^{\mathrm{e}}$&
$-8.2\pm1.1^{\mathrm{d}}$&
\ldots{}&
no\tabularnewline
\hline
\end{tabular}\par\end{center}

\end{table*}

%==========================================

\section{Results}

\label{sec:Results}

\subsection{Radial velocities}

\label{sub:Radial-velocities}

%==========================================

\begin{table*}

\caption{Summary of the observations, the heliocentric radial velocities ($\mathrm{RV}$)
from two-epoch and the average rotational velocities ($v\sin i$).
The uncertainties of relative radial velocities ($\sigma_{\mathrm{RRV}}$)
with respect to the template star LHS~049 and the average radial
velocities ($\overline{\mathrm{RV}}$) are also given. }

\label{tab:summary_results}

\begin{center}\begin{tabular}{lcrrrrr}
\hline 
0.81&
Date&
HJD-2453100&
RV&
$\sigma_{\mathrm{RRV}}$&
$\overline{\mathrm{RV}}$&
$v\sin i$\tabularnewline
&
&
&
$\left[\mathrm{km\, s^{-1}}\right]$&
$\left[\mathrm{km\, s^{-1}}\right]$&
$\left[\mathrm{km\, s^{-1}}\right]$&
$\left[\mathrm{km\, s^{-1}}\right]$\tabularnewline
\hline 
GY 5&
2004-Apr-24&
20.7198440&
$-6.77\pm2.02$&
1.96&
&
\tabularnewline
&
2004-May-07&
33.7721938&
$-6.02\pm0.85$&
0.68&
$-6.39\pm2.19$&
$16.5\pm0.6$\tabularnewline
GY 141&
2004-May-10&
36.6712108&
$-4.15\pm0.63$&
0.38&
&
\tabularnewline
&
2004-May-17&
43.6683007&
$-3.46\pm0.66$&
0.43&
$-3.81\pm0.91$&
$4.4\pm1.4$\tabularnewline
GY 310&
2004-Apr-24&
20.8373504&
$-5.27\pm0.92$&
0.77&
&
\tabularnewline
&
2004-May-09&
35.8224052&
$-8.19\pm0.71$&
0.51&
$-6.73\pm1.16$&
$11.1\pm6.0$\tabularnewline
USco 40&
2004-Apr-05&
1.7661754&
$-7.34\pm0.52$&
0.15&
&
\tabularnewline
&
2004-May-07&
33.7620554&
$-6.28\pm0.52$&
0.15&
$-6.81\pm0.74$&
$34.2\pm0.5$\tabularnewline
USco 53&
2004-Apr-04&
0.9036653&
$-8.27\pm0.70$&
0.50&
&
\tabularnewline
&
2004-May-02&
28.7394676&
$-7.22\pm0.63$&
0.39&
$-7.75\pm0.95$&
$40.0\pm0.6$\tabularnewline
USco 55&
2004-Apr-05&
1.8422807&
$-6.21\pm0.50$&
0.04&
&
\tabularnewline
&
2004-May-02&
28.8141198&
$-6.55\pm0.50$&
0.06&
$-6.38\pm0.71$&
$22.9\pm0.8$\tabularnewline
USco 66&
2004-Apr-05&
1.7972634&
$-7.24\pm0.95$&
0.81&
&
\tabularnewline
&
2004-May-02&
28.7956003&
$-8.38\pm1.32$&
1.22&
$-7.81\pm1.63$&
$25.9\pm1.2$\tabularnewline
USco 67&
2004-Apr-05&
1.7188620&
$-7.07\pm0.67$&
0.44&
&
\tabularnewline
&
2004-May-02&
28.7113799&
$-5.70\pm0.58$&
0.30&
$-6.38\pm0.88$&
$18.4\pm0.4$\tabularnewline
USco 75&
2004-Apr-04&
0.8840376&
$-6.80\pm0.70$&
0.50&
&
\tabularnewline
&
2004-May-07&
33.6065432&
$-6.16\pm0.78$&
0.60&
$-6.48\pm1.05$&
$55.6\pm3.0$\tabularnewline
USco 100&
2004-Apr-06&
1.8179138&
$-7.37\pm0.67$&
0.44&
&
\tabularnewline
&
2004-May-02&
28.7752928&
$-7.47\pm0.62$&
0.36&
$-7.42\pm0.91$&
$43.7\pm3.2$\tabularnewline
USco 101&
2004-Apr-04&
0.8120734&
$-3.25\pm0.66$&
0.44&
&
\tabularnewline
&
2004-May-02&
28.6591660&
$-5.03\pm0.60$&
0.32&
$-4.14\pm0.89$&
$19.1\pm0.3$\tabularnewline
USco 104&
2004-Apr-04&
0.7850385&
$-7.05\pm0.61$&
0.35&
&
\tabularnewline
&
2004-May-02&
28.6349480&
$-7.62\pm0.57$&
0.27&
$-7.33\pm0.83$&
$16.7\pm0.4$\tabularnewline
USco 109&
2004-Apr-05&
1.7453989&
$-4.72\pm0.50$&
0.07&
&
\tabularnewline
&
2004-May-07&
33.6304878&
$-4.70\pm0.51$&
0.09&
$-4.71\pm0.72$&
$8.6\pm1.2$\tabularnewline
USco 112&
2004-Apr-04&
0.8552168&
$-2.95\pm0.50$&
0.05&
&
\tabularnewline
&
2004-May-07&
33.5826368&
$-3.41\pm0.51$&
0.07&
$-3.18\pm0.71$&
$5.8\pm1.2$\tabularnewline
USco 121&
2004-Apr-24&
20.7059408&
$-40.76\pm0.90$&
0.74&
&
\tabularnewline
&
2004-May-02&
28.6969641&
$-42.48\pm0.54$&
0.20&
$-41.62\pm1.04$&
$17.6\pm1.3$\tabularnewline
USco 128&
2004-May-13&
39.7978276&
$-7.03\pm0.51$&
0.09&
&
\tabularnewline
&
2004-May-17&
43.6108024&
$-6.32\pm0.51$&
0.10&
$-6.68\pm0.72$&
$3.6\pm1.1$\tabularnewline
USco 130&
2004-May-09&
35.7724090&
$-3.92\pm0.94$&
0.80&
&
\tabularnewline
&
2004-May-13&
39.8538830&
$-3.90\pm0.67$&
0.45&
$-3.91\pm1.16$&
$15.2\pm1.1$\tabularnewline
USco 132&
2004-May-13&
39.8268683&
$-6.68\pm0.53$&
0.18&
&
\tabularnewline
&
2004-May-17&
43.6391138&
$-6.61\pm0.54$&
0.20&
$-6.64\pm0.76$&
$9.1\pm0.7$\tabularnewline
\hline
\end{tabular}\par\end{center}

\end{table*}

%==========================================

%==========================================

\begin{figure*}

\begin{center}\includegraphics[clip,scale=1.3]{figures/radvel_all01}\par\end{center}

\caption{Relative radial velocities (RVs) of objects measured in two different
epochs. The vertical axes indicate the amount of deviation ($\Delta\mathrm{RV}$)
from the `average' radial velocity ($\overline{\mathrm{RV}}$) in
Table~\ref{tab:summary_results}), and the horizontal axes indicate
the time of the observation in heliocentric Julian date (HJD). The
objects are considered to have a non-constant RV when the error bars
of two data points do not overlap each other. The non-constant RV
objects are considered as binary/multiple candidates.}

\label{fig:radvel01}

\end{figure*}

%==========================================

\addtocounter{figure}{-1}

%==========================================

\begin{figure*}

\begin{center}\includegraphics[clip,scale=1.3]{figures/radvel_all02}\par\end{center}

\caption{continued}

\label{fig:radvel02}

\end{figure*}

%==========================================

The radial velocities of each object were determined by using the
cross-correlation function of the object spectrum with that of a template
star which has a similar spectra type. By visual inspection, the wavelength
ranges used for the cross-correlation calculations are chosen by avoiding
the regions of spectra affected by the telluric lines, and defects
and fringes (in near infrared) of the CCDs. The radial velocities
of objects with respect to the template are obtained by measuring
the location of the peak in the cross-correlation function. The location
of the peak is determined by fitting the section of the cross-correlation
function around the peak by a second order polynomial. LHS~49 (Proxima
Cen, M5.5) was chosen as the template for this purpose. The heliocentric
radial velocity of the template object LHS~49 was obtained by measuring
the wavelength shifts of the prominent photospheric absorption features
\ion{K}{i}~$\lambda\lambda$7664.911,~7698.974. This gives us
$\mathrm{RV_{LHS49}=}-22.6\pm0.5\,\kmps$, which is in good agreement
with the earlier measurement of \citet{garcia-sanchez:2001} who found
$\mathrm{RV_{LHS49}=}-21.7\pm1.8\,\kmps$. The heliocentric RV of
each object can be then calculated by adding $\mathrm{RV_{LHS49}}$
with the RV of each object with respect to LHS~49. In the following
measurements of the heliocentric radial velocities, our measurement
($\mathrm{RV_{LHS49}=}-22.6\pm0.5\,\kmps$) will be used for consistency.

Before applying the cross-correlation technique to our main targets,
we have applied the technique to the radial velocity standard HD~140538
for which an high-accuracy RV measurement via the fixed-configuration,
cross-dispersed échelle spectrograph Elodie \citep{baranne:1996}
is available. This was done so to ensure not only the validity of
the cross-correlation technique, but also the validity of the wavelength
calibration. In this test, we found the heliocentric $\mathrm{RV_{HD\,140538}=}18.18\pm0.62\,\kmps$
which is in good agreement with the Elodie radial velocity measurement
of $19.00\pm0.05\,\mathrm{km\, s^{-1}}$ \citep{udry:1999}.

The result of the heliocentric RV measurements (from two epochs for
each object) is summarised in Table~\ref{tab:summary_results} along
with the uncertainties. The table also lists the uncertainties in
the `relative' radial velocities ($\sigma_{\mathrm{RRV}}$) which
is the uncertainty of RV with respect to the template star. To determine
the $\sigma_{\mathrm{RRV}}$, we have used the following procedure:
(1)~add random gaussian deviate noise to the object spectra using
the corresponding variance spectra, (2)~compute the cross-correlation
curve using the spectra with added noise and the template, (3)~measure
the RV by location the peak, (4)~repeat 1--3 for 100 times, and computes
the standard deviation of the 100 RV values. This should provide a
good estimate of uncertainties in RV cased by the uncertainties in
the flux levels. %The two consecutive measurements of RVs (from same nights) are used to
%find the (absolute/heliocentric) RV of the night, and their uncertainties are obtained from
%the standard deviations of the mean ($\sigma_{\mathrm{RRV}}$) and
%the uncertainty of the template RV value (c.f.~\citealt{joergens:2006}).
Note that for finding a shift in the radial velocities from two different
epochs, only $\sigma_{\mathrm{RRV}}$ is important since one only
requires relative (with respect to a template) RV values. The `average'
radial velocities $\left(\overline{\mathrm{RV}}\right)$ of the two
epochs are also given in the same table. The $\overline{\mathrm{RV}}$
values of the objects in UScoOB are consistent with the earlier measurements
of \citet{muzerolle:2003} in Table~\ref{tab:literatures}.

For each object and for each RV measurement, the deviations ($\Delta\mathrm{RV}$)
from the average RV are computed and summarized in Fig.~\ref{fig:radvel01}
along with their uncertainties ($\sigma_{\mathrm{RRV}}$) in order
to aid the identification of multiplicity candidates. Note that in
computing $\Delta\mathrm{RV}$ we do not require the knowledge of
absolute radial velocities, but only the relative velocities (with
respect to a template). Alternatively, we also computed the cross-correlation
functions of spectra of a same object (instead of using the spectra
of the template LHS~49) from two different epochs in order to find
$\Delta\mathrm{RV}$. The resulting cross-correlation functions were
found to be too noisy (due to the relatively low S/N in the object
spectra), and the peak positions of the cross-correlation curves were
not reliably located.

To identify an object with a RV variation with a statistical significance
from our sample, we apply the method described by \citet{maxted:2005},
which we briefly summarize next, to our data. There are three steps
in this method: (1)~compute the $\chi^{2}$ by fitting the two-epoch
RV data for each object with a constant function (a zero-th order
polynomial), (2)~compute the corresponding $\chi^{2}$ probability
($p$), (3)~designate the object as a non-constant RV object or a
binary candidate if $p<10^{-3}$ (1~per~cent). When computing $\chi^{2}$,
we use the uncertainties in the relative RV ($\sigma_{\mathrm{RRV}}$).
The number of degree of freedom in the fitting procedure is obviously
1. A similar method was also used in a recent RV survey of VLM stars
by \citet{basri:2006}.

We have computed $p$ for all the objects, and have plotted the results
as a histogram of the $-\log{p}$, shown in Fig.~\ref{fig:chisq-test}
(excluding the non-member USco~121; see explanation later). The figure
clearly shows two distinctive populations: one on the left (with small
$-\log{p}$ values) occupied by the RV constant objects, and one on
in the right-most bin (with the largest $-\log{p}$) occupied by the
RV variable objects which have $p<10^{-3}$. The expected distribution
of $p$ computed consistently with our uncertainty measurements is
also shown in the same figure. The RV constant population on the left
side reasonably matches the expected curve. This reassures that our
uncertainties in RV values are reasonable, and an additional systematic
error may not be necessary in this analysis (but see Section~\ref{sec:Binary-fraction}).
According to the figure, there are four objects which show significant
RV variations out of 17 samples. They are GY~141, GY~310, USco~55,
and USco~104, and considered as our preliminary binary/multiple candidates.
We will discuss the effect of the systematics error in detail, and
list our final binary/multiple candidates in Section~\ref{sec:Binary-fraction}.

%==========================================

\begin{figure}

\begin{center}\includegraphics[clip,scale=0.63]{figures/chisq_hist} \par\end{center}

\caption{Histogram of the $\chi^{2}$ probability ($p$) for fitting the observed
(relative) RV values with a constant (a horizontal line). The constant
used in the fit is determined from the weighted mean of the two RV
measurements for each object. The objects with $p<10^{-3}$ are identified
as \emph{non-constant} or \emph{multiple}, which appear in the right-most
bin in the histogram. There are four objects which satisfy this condition.
The expected distribution (dashed) for RV constant objects based on
the RV uncertainties in the observations (Table~\ref{tab:summary_results})
and based on the $\chi^{2}$ probability function is also shown for
a comparison. The match between the expected distribution and the
histogram is reasonable, indicating that our uncertainty estimates
in RV values are also reasonable. }

\label{fig:chisq-test}

\end{figure}

%==========================================

Finally, the histogram of $\overline{\mathrm{RV}}$ for the objects
in UScoOB is given in Fig.~\ref{fig:vrad_histrogram}. The total
number of the objects is 14. Note that USco~121 is excluded from
the graph since it is identified as a non-member of the UScoOB association
based on the RV value (see Table~\ref{tab:summary_results}). \citet{muzerolle:2003}
also found it to be a likely non-member based on the radial velocity
and the low lithium abundance. The distribution of the RVs in the
figure was fitted by a gaussian function. We found that the standard
deviation and the peak position of the radial velocity distribution
are $1.2\,\kmps$ and $-5.9\,\kmps$ respectively. The former is very
similar to the standard deviation (0.9~$\kmps$) of the radial velocity
distribution of 9 BDs and VLM objects in Cha~I found by \citet{joergens:2006b}.
They also studied the radial velocity distribution of more massive
25 T~Tauri stars in Cha~I, and found the standard deviations ($1.3\,\kmps$)
is not significantly different from that of the brown dwarfs and the
very low-mass objects. Unfortunately, we do not have the radial velocity
measurements of the higher mass counter parts (T~Tauri stars) in
Upper Sco OB association. This is planned for a near future investigation
as this is important for the study of the mass dependency of the kinematics
in a young stellar cluster.

According to the hydrodynamical simulations of a low-mass star-forming
cluster of \cite{bate:2003} which yields a stellar density of $\sim10^{3}\,\mathrm{stars\, pc^{-3}}$,
the rms dispersion (1-D) of the stars and the BDs is $1.2\,\kmps$.
Similarly for the model with a higher stellar density ($\sim10^{4}\,\mathrm{stars\, pc^{-3}}$),
the rms dispersion is 2.5~$\kmps$ (\citealt{bate:2005}). The standard
deviation of $\overline{\mathrm{RV}}$ ($1.2\,\kmps$) found in our
analysis is more comparable the lower stellar density model.

%==========================================

\begin{figure}

\begin{center}\includegraphics[clip,scale=0.63]{figures/radvel_hist}\par\end{center}

\caption{Histogram of the average heliocentric radial velocities of 14 UScoOB
BD and VLM objects listed in Table~\ref{tab:summary_results} (excluding
USco~121, a non-member). The gaussian fit (dashed) of the radial
velociity distribution gives a standard deviation of $1.2\,\kmps$
and the peak position of $-5.9\,\kmps$. }

\label{fig:vrad_histrogram}

\end{figure}

%==========================================

\subsection{Rotational velocities}

\label{sub:Rotational-velocities}

The rotational velocities of the objects were determined by measuring
the widths of the cross correlation functions of the target spectra
against a template spectrum from an object which is known to have
a very small rotational velocity. The line broadening of the targets
is assumed to be dominated by rotational broadening. As in the cases
for the radial velocity measurements, LHS~49 is chosen as the template.
Using its rotational period ($P\approx83\,\mathrm{d}$, \citealt{benedict:1998})
and radius ($R_{*}\approx0.145\,\mathrm{R_{\odot}}$ from the VLTI
measurement by \citealt{segransan:2003}), the rotational velocity
of LHS~49 is estimated as $v\,\sin i=2\pi R_{*}/P\approx0.1\,\kmps$;
negligibly small.

The width of the cross-correlation curves ($\sigma_{\mathrm{CCF}}$)
are calibrated with the rotational velocities ($v\,\sin i$) by cross
correlating the template spectra against the same template spectra
with added rotation (convolved with a given $v\,\sin i$), as done
by e.g.~\citet{tinney:1998}, \citet{mohanty:2003} and \citet{white:2003}.
A linear limb-darkening law with a solar-like parameter ($\epsilon=0.6$)
was assumed in the formulation of the rotational profile described
by \citet{gray:1992}, his Eq.~17.12. For each object, two measurements
of rotational velocities are computed from two independent spectra
obtained at different epochs. As for the RV measurements, the mean
and the standard deviation of the mean are used as the final rotational
velocity and its uncertainty. The final results are recorded in Table~\ref{tab:summary_results}.
In general, our measurements are in good agreement with the earlier
measurements of \citet{muzerolle:2003} and \citet{mohanty:2005},
given in Table~\ref{tab:literatures}. For example, \citet{muzerolle:2003}
found $v\,\sin i=16.8\pm2.7\,\mathrm{km\, s^{-1}}$ for GY~5 while
we found $v\,\sin i=16.5\pm0.6\,\mathrm{km\, s^{-1}}$.

The range of $v\,\sin i$ found among our objects is $3.6$--$55.6\,\kmps$,
and a similar range is also found by \citet{mohanty:2005}. Fig.~\ref{fig:vrot_histrogram}
shows the histogram of $v\,\sin i$ distribution for the UScoOB objects
(14 objects excluding USco 121, non member). The log-normal fit of
this distribution gives the peak position at $16.9\,\kmps$ with a
standard deviation $\sigma=27.8\,\kmps$. Using the $v\,\sin i$ data
in \citet{joergens:2001}, the same histogram bin size used for UScoOB
obects and a log-normal fit, we find the $v\,\sin i$ distribution
of the BD and VLM stars (8 objects) in Cha~I peaks at $15.4\,\kmps$,
and has the standard deviation of $8.0\,\kmps$. The peak of the distribution
is similar to that of Upper Sco objects, but the standard deviation
of the distribution is much smaller than that of our Upper Sco objects.
The difference maybe due to the very small sample. A similar fit was
applied to the $v\,\sin i$ distribution of 14 higher mass T~Tauri
stars in Cha~I using the data of \citet{joergens:2001}, and we found
a peak at $17.0\,\kmps$ with a standard deviation $25.9\,\kmps$
which are very similar to those of the Upper Sco brown dwarf candidates
and VLM stars.

%==========================================
\begin{figure}

\begin{center}\includegraphics[clip,scale=0.63]{figures/vrot_hist}\par\end{center}

\caption{Histogram of the rotational velocities ($v\,\sin{i}$) of 14 UScoOB
BD and VLM objects listed in Table~\ref{tab:summary_results} (excluding
USco~121, a non-member). The log-normal fit (dashed) of the rotational
velocity distribution gives a standard deviation $27.8\,\kmps$ and
the peak poistion $16.9\,\kmps$.}

\label{fig:vrot_histrogram}

\end{figure} %==========================================

\section{Binary fraction}

\label{sec:Binary-fraction}

In Section~\ref{sub:Radial-velocities}, we found 4 out of 17 (excluding
USco~121; non-member) objects show a statistically significant RV
variation, indicating that they are binary/multiple candidates. In
order to estimate the uncertainty in the binary/multiple fraction
from this relatively small sample, we will follow the method used
by \citet{basri:2006} who considered the binomial distribution, $P_{B}\left(x,n,p\right)$
of $x$ positive event out of $n$ trials with the probability $p$
for a positive event in each trial. In our case, $n=17$ (the number
of sample) and $x=4$. The peak of $P_{B}\left(4,17,p\right)$ curve
suggests the binary fraction of 24~per~cent, as its should be ($4/17=0.24$).
The uncertainties was estimated by plotting $P_{B}\left(4,17,p\right)$,
and finding the values of $p$ at which $P_{B}$ reduced to $e^{-1}$
of the peak value. In this analysis, we find the binary/multiple fraction
along with the uncertainties of our sample to be $f=24_{-13}^{+16}$~per~cent.

Next, we investigate the range of semimajor axes of binaries (equivalently
the range of binary periods) to which our RV survey is sensitive.
For this purpose, we will consider the detection probability for binaries
or the RV variables, given the time separations of two-epoch observations
and the ranges of estimated primary masses (c.f.~Tables~\ref{tab:literatures}
and \ref{tab:summary_results}). The probability is calculated based
on the simulated RV observations of binaries whose orbit are randomly
selected from a model. A similar method was used by \citet{maxted:2005}
and \citet{basri:2006}. The most important factors in determining
the detection probability are the size of uncertainties in RV measurements
(which we use the average $\sigma_{\mathrm{RRV}}$ from our observation),
and the time separations of observations. The smaller the errors in
RV measurements, the larger the probability for a given binary orbit
and a time separation of orbit. The larger the time separation of
observations, the larger the upper limit of the semimajor axis to
which an observation is sensitive. The average time separation of
the two-epoch RV observations of our targets is $21.6$~d. In the
following, we will briefly discuss our model assumptions and parameters
which are essentially the same as those of \citet{maxted:2005} but
with some simplifications.

There are six basic parameters in our Monte Carlo simulation: primary
mass ($M_{1}$), mass ratio ($q$), eccentricity ($e$), orbital phase
($\phi$), orbital inclination ($i$), longitude of periastron ($\lambda$).
The primary mass $M_{1}$ is assigned from the adopted mass of the
targets in Table~\ref{tab:literatures}, with a uniform random deviation
of $\pm0.002\, M_{\sun}$. The mass ratio is assumed to be uniformly
distributed between $q=0$ and $0.2$. The eccentricity $e$ is assumed
to be zero (circular orbits). Both \citet{maxted:2005} and \citet{basri:2006}
found the detection probability is insensitive to the assumed distribution
of $q$ and $e$. The orbital phase is randomly chosen between 0 and
1. The inclination $i$ is randomly chosen from the cumulative distribution
of $\cos{i}$. The longitude of periastron $\lambda$ is not necessary
since we assumed $e=0$.

In order to compute the detection probability as a function of semimajor
axis $a$, we take the following procedure: (1)~for each object in
our targets, we randomly select $10^{5}$ binaries using the assumption
stated above for a given value of $a$, and compute the RV of the
primary ($V_{1}$), (2)~evolve the orbit by the time separation of
the RV measuremets used in the observations for this object, and take
another simulated measurement of RV ($V_{2}$), (3)~from $V_{1}$,
$V_{2}$ and the average uncertainty in RV ($\sigma_{\mathrm{RRV}}$)
from the observations, we compute the $\chi^{2}$ probability $p$,
and flag the trial as a detection if $p<10^{-3}$ as done for the
real data, (4)~find the fraction of detections out of all random
trials, (5) repeat 1--4 for the range of $a$ between $10^{-3}$~and
10~au, and (6) repeat 5 for the all targets, and find the detection
probability averaged over all targets as a function of $a$.

The result of the simulation is shown in Fig.~\ref{fig:detect_prob}.
The detection probability curve shown here is very similar that of
\citet{basri:2006} for a constant time separation (20~d) case (see
their Fig.~4). The probability remains fairly constant up to $a\approx0.1\,\mathrm{au}$,
and it rapidly decreases beyond $a\approx0.3\,\mathrm{au}$. This
turning point will increase if we had used larger time separations
in our observation. From this figure, we find that the 80~per~cent
detection probability up to for binaries with $a<0.1\,\mathrm{au}$,
given our time separations and the average uncertainty ($\sim0.56\,\kmps$)
in the RV measurements.

%==========================================
\begin{figure}

\begin{center}\includegraphics[clip,scale=0.63]{figures/detect_prob} \par\end{center}

\caption{The binary detection probability (solid) as a function of the semimajor
axis ($a$), based on Monte~Carlo simulations of RV measurements,
is shown. For each object in our target list, the simulation was performed
using the time separation ($\Delta t$) actually used in our two-epoch
observations (c.f.~Table~\ref{tab:summary_results}) and the estimated
primary mass (c.f.~Table~\ref{tab:literatures}). The final detection
probability is obtained by averaging over the simulated observations
of all the objects. The average time separation in the two-epoch observations
is $\sim21.6\,\mathrm{d}$, and the average uncertainty in RV is $~0.56\,\kmps$.
The probability remains fairly constant up to $a\approx0.1\,\mathrm{au}$,
and it rapidly decreases beyond $a\approx0.3\,\mathrm{au}$. The 80~per~cent
detection probability is achieved for binaries with $a<0.1\,\mathrm{au}$. }

\label{fig:detect_prob}

\end{figure}

%==========================================

%% The expected binary fraction for our survey can be obtained
%% by integrating the product of the detection probability found above
%% with an assumed binary separation distribution function over a range
%% of $a$. Here, we simply adopt one of the separation distribution
%% function used by \citet{maxted:2005}. The binary separation ($a$)
%% distribution is assumed to be  gaussian in  log($a$) space with its
%% centre $\log{a_{o}}=0.6$ and width $\sigma_{\log{a}}=1.0$, and
%% truncated at $\log{a}=1.176$ where $a$ is in au. This form of the
%% distribution function is motivated by the distribution of solar-type
%% binaries found in \citet{duquennoy:1991}, and the truncation is
%% suggested from the lack of wide binaries for $a>15$~au 
%% (e.g.~\citealt{burgasser:2003}). The distribution is normalized such
%% that the total binary fraction of 26~per~cent as
%% suggested by \citet{basri:2006}.  The resulting separation
%% distribution function is shown also in Fig.~\ref{fig:detect_prob}. 
%% Using this and the detection probability in the same figure, we find
%% the expected binary fraction to be $f_{\mathrm{exp}}=3.2$~per~cent
%% (for $a<0.1$~au), which is
%% much smaller than our observed binary fraction of
%% $f_{\mathrm{obs}}=24^{+16}_{-13}$. The two values are not consistent
%% with each other. There are two possible reasons for the discrepancy:
%% (1)~the assumed binary separation distribution mentioned above is
%% unrealistic, and (2)~the uncertainties in our RV measurements are
%% underestimated.  The first reason is less likely since both
%% \citet{maxted:2005} and \citet{basri:2006} found this form of the $a$
%% distribution function is consistent with the observed binary
%% fractions estimated from many earlier surveys.  We will investigate
%% the possibility for the second reason in the following.

In Section~\ref{sub:Radial-velocities} (see Fig.~\ref{fig:chisq-test}),
we found the histogram of the $\chi^{2}$ probability $p$, for the
RV data fit assuming constant RV, reasonably matches the expected
distribution. This leas us to conclude that the uncertainty estimates
in RV values are reasonable. Here, we will inspect the effect of the
systematic error on the observed binary fraction more closely. There
are two possible source of errors: (1)~the RV shifts caused by surface/atmospheric
activity, and (2)~the RV shifts due to the limit of the stability
of autoguiding of the telescope. \citet{joergens:2006} found that
(1) is negligibly small for young VLM objects with $M_{*}<0.12\, M_{\sun}$.
The size of RV errors caused by (2) is expected to be $<1\,\kmps$
\citep{bailer-jones:2004}. Note that \citet{maxted:2005} found an
additional systematic error was not needed to be added to the RV uncertainties
of \citet{joergens:2006} who used the same instruments as ours, and
used the same method of estimating the uncertainties. Hence, we expect
the systematic error to be much smaller than $1\,\kmps$ if one is
required.

The observed binary fraction is expected to be sensitive to the assumed
size of the systematic error ($\sigma_{\mathrm{sys}}$) since the
RV shifts and the systematic errors are in the same order of magnitude.
The number of binary detections would decrease as $\sigma_{\mathrm{sys}}$
increases. We have computed the $\chi^{2}$ probabilities ($p$) for
each object being a RV constant object as done earlier (Section\ref{sub:Radial-velocities}),
but with the assumed systematics error of $\sigma_{\mathrm{sys}}=0.2\,\kmps$,
which was added to the uncertainties in the RVs of each object. The
results are summarised in Table~\ref{tab:prob_compare} along with
the probabilities computed without the systematic error (as in Section\ref{sub:Radial-velocities}).
The number of objects with $p<10^{-3}$ decreases from 4 to 2 as the
systematic error is introduced. Unfortunately, the probabilities computed
with $\sigma_{\mathrm{sys}}=0.0\,\kmps$ for GY~141 and USco~55
are just below the $p=10^{-3}$ boundary value; hence, they become
RV constant objects as the systematic error is introduced. This clearly
demonstrates the importance of the uncertainty estimates in the measured
RV values. We found that the histogram of the $\chi^{2}$ probabilities
with $\sigma_{\mathrm{sys}}=0.2\,\kmps$ (as in Fig~\ref{fig:chisq-test}),
is also in reasonable agreement with the expected distribution of
$-\log{p}$ (not shown here). The binary fraction estimated with the
assumed systematic error is $2/17$ or $f=12_{-8}^{+13}$~per~cent
for $a<0.1$~au.

% This may be out of place =======================
An alternative explanation for the RV variations found in the binary
candidates above is a type of stellar pulsations found by \citet{palla:2005}.
They studied the non-adiabatic, linear instability of very VLM stars
and BD during the deturium burning phase in the core, and found unstable
fundamental modes in the time-scale between $\sim1$ and $\sim5$~h
for the object mass from $0.02$ and $0.1\,\Msun$. By using their
pulsation periods and by assuming that the pulsation amplitudes is
5~per~cent of the radius of the objects, we find that the expected
RV variation can be $>6\,\kmps$. As mentioned in Section~\ref{sec:observation-reduction},
for each object at a given night, two separate spectra are obtained
consecutively (with time separations of $\sim0.5$~h.). Before co-adding
the two spectra together, we have measured the RV from each spectrum
to check if there is a large jump in the RV measurements in short
time-scale. The shifts in the RV values in measured in the two consecutive
spectra are found to be between $0.02\pm0.10$ (USco~128) and $2.04\pm1.67\,\kmps$
(USco~66), which is much smaller than the expected RV variation based
on the models of \citet{palla:2005}; hence, found little indication
of the pulsations.

%% %==========================================
%% \begin{figure}

%% \begin{center}

%% \includegraphics[%
%%   clip,
%%   scale=0.63]{figures/detect_err_add.eps}

%% \end{center}

%% \caption{The number of binaries/multiples (non-constant RV objects)
%%   detected as a function of the size of additional systematic error ($\sigma_{\mathrm{sys}}$)
%%   in RV measurements. The number of detections decreases rapidly from 4
%%   to 2 as $\sigma_{\mathrm{sys}}=0.1\,\kmps$, which may be caused by
%%   the instability of the autoguiding of the telescope, is added to
%%   the uncertainties estimated in
%%   Table~\ref{tab:summary_results}. The number of  detections
%%   remains constant at 2 for a relatively wide range of the additional
%%   errors. i.e.~$0.05\,\kmps < \sigma_{\mathrm{sys}} < 0.275\,
%%   \kmps$. No binary is found for  $\sigma_{\mathrm{sys}} > 0.375\,\kmps$.} 
%% \label{fig:adding_sys_err}

%% \end{figure}
%% %==========================================

%==========================================
\begin{table}

\caption{The $\chi^{2}$ probability ($p$), for each object being a RV constant,
listed in ascending order of $p$. The ellipses represent the boundary
($p=10^{-3}$) between RV constant objects and RV variable objects. }

\label{tab:prob_compare}

\begin{center}\begin{tabular}{lc}
\hline 
obj.~ID&
$p$\tabularnewline
\hline 
USco~112&
$8.9\times10^{-8}$\tabularnewline
USco~128&
$1.3\times10^{-7}$\tabularnewline
USco~40&
$5.8\times10^{-7}$\tabularnewline
USco~55&
$2.4\times10^{-6}$\tabularnewline
$\cdots$&
$\cdots$\tabularnewline
USco~101&
$1.1\times10^{-3}$\tabularnewline
GY~310&
$1.6\times10^{-3}$\tabularnewline
USco~67&
$1.0\times10^{-2}$\tabularnewline
USco~53&
$9.8\times10^{-2}$\tabularnewline
USco~104&
$2.0\times10^{-1}$\tabularnewline
GY~141&
$2.3\times10^{-1}$\tabularnewline
USco~75&
$4.1\times10^{-1}$\tabularnewline
USco~66&
$4.4\times10^{-1}$\tabularnewline
GY~5&
$7.2\times10^{-1}$\tabularnewline
USco~132&
$7.9\times10^{-1}$\tabularnewline
USco~100&
$8.6\times10^{-1}$\tabularnewline
USco~109&
$8.6\times10^{-1}$\tabularnewline
USco~130&
$9.8\times10^{-1}$\tabularnewline
\hline
\end{tabular}\par\end{center}

\end{table} %==========================================

%% %==========================================
%% \begin{table}

%% \caption{Summary of the observed binary fractions ($f_{\mathrm{obs}}$) compared with the
%%   expected binary fractions ($f_{\mathrm{exp}}$) with and without a
%%   systematic error ($\sigma_{\mathrm{sys}}$). }

%% \label{tab:binary_fraction}

%% \begin{center}

%% \begin{tabular}{ccccl}
%% \hline 
%% $\sigma_{\mathrm{sys}}$&  
%% N. detected&  
%% $f_{\mathrm{obs}}$&
%% $f_{\mathrm{exp}}$&  
%% obj.~ID  \tabularnewline
%% ($\kmps$)&
%% &
%% (per~cent)&
%% (per~cent)&
%% \tabularnewline
%% \hline
%% $0.0$&
%% $4$&
%% $24^{+16}_{-13}$&
%% $3.2$&
%% GY~141\tabularnewline
%% &
%% &
%% &
%% &
%% GY~310\tabularnewline
%% &
%% &
%% &
%% &
%% USco~55\tabularnewline
%% &
%% &
%% &
%% &
%% USco~104\tabularnewline
%% $0.1$&
%% $2$&
%% $12^{+13}_{-10}$&
%% $2.9$&
%% GY~310\tabularnewline
%% &
%% &
%% &
%% &
%% USco~104\tabularnewline
%% \hline
%% \end{tabular}

%% \end{center}

%% \end{table}
%==========================================

\section{Conclusions}

\label{sec:Conclusions}

We have presented two-epoch RV survey of 18 young BDs and VLM stars
($0.02~M_{\sun}<M_{*}<0.1~M_{\sun}$) in UScoOB and $\rho$~Oph dark
cloud core using the high resolution UVES echelle spectroscopy at
VLT. The average time separation of RV measurements are 21.6~d, and
our RV survey is sensitive to binaries with separation smaller than
0.1~au. One of our targets, USco~121, is most likely a non-member
of the UScoOB association based on the deviation of the RV from the
result of the population in the association. A similar conclusion
was found by \citet{muzerolle:2003} from their RV study and the low
lithium abundance.

We found 4 (GY~141, GY~310, USco~55 and USco~104) out of 17 objects
as our binary/multiple candidates. This corresponds to the binary
fraction of $24_{-13}^{+16}$~per~cent for the binary separation
$a<0.1$~au. We found that the estimated binary fraction is very
sensitive to the assumed size of systematic error. If an additional
systematic error of $0.2\,\kmps$ is added to the uncertainties in
the RV measurements, we find only 2 (GY~310 and USco~104) out of
17 objects as our binary/multiple candidates, which corresponds to
the binary fraction of $12_{-10}^{+13}$~per~cent.

The recent high-resolution imaging survey of brown dwarfs and very
low-mass objects (M5.5--M7.5 ) in the UScoOB by \citet{kraus:2005}
(which was not known to authors at the time of our observation: April--May,
2004) confirms that USco~55 and USco~66 are multiple systems, and
USco~109 is most likely a multiple system. Their projected separations
are $a>4.0$~au which is well above our sensitivity limit of 0.1~au;
therefore, it is not surprising that we did not identify USco~66
and USco~109 as multiples. Interestingly, we also found USco~55
as a candidate for a multiple system. This may indicate that USco~55
is a triple system with one of object located within 0.1~au from
the primary. In addition, we identify USco~104 as a binary candidate,
but this was not in their target list. Further they found USco~67,
75, 112, 128, 130, 132 as non-multiple, and similarly we found them
as 'non-candidates' for multiplicity.

We found the RV dispersion ($1.2\,\kmps$) of the objects in UScoOB
is very similar to that of the BDs and VLM stars in Chamaeleon~I
(Cha~I) previous study by \citet{joergens:2006b}. The rotational
velocities ($v\,\sin i$) of the samples were also measured. The distribution
of $v\,\sin i$ for the UScoOB objects peaks around $16.9\,\kmps$
which is also similar to that of the Cha~I population found by \citet{joergens:2006b};
however, the dispersion of $v\,\sin i$ for the UScoOB objects ($27.8\,\kmps$)
is found to be much larger than that of the Cha~I objects ($8.0\,\kmps$)

Follow-up spectroscopic observations of the binary candidates presented
here are planned in near future. There are only a few RV variable
binary candidates identified in earlier surveys (\citealt{guenther:2003};
\citealt{kenyon:2005}; \citealt{joergens:2006}). As \citet{burgasser:2006}
points out most of the current RV and imaging surveys use samples
from magnitude-limited survey, but one should attempt to use the samples
from volume-limited survey in order to find a correct statistics on
binary parameters more straightforwardly, i.e. without correcting
for bias.

\section{{*}}

{Acknowledgements}

We thank the staff of VLT of the ESO for carrying out the observations
in service mode. RK is grateful for Rob Jeffries for helpful suggestions
on the data analysis presented in this paper. RK also thank Matteww
Bate and Tim Naylor for providing us helpful comments on the manuscript.
This work was supported by PPARC rolling grant PP/C501609/1.

%----------------------------------------------------

% Bibliography follows here.

%-----------------------------------------------------

\bibliographystyle{mn}

\bibliographystyle{mn}
\bibliography{local}

\label{lastpage} 
\end{document}